---
title: "Long-term WQ: Sampling Effort from DroughtData"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Look at the sampling effort for the various parameters to be included in the long-term WQ publication. This will help us decide how to structure our ANOVA models. This document also contains the code used to create a map and a summary table of the surveys, parameters, and periods of record for the publication.


# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(scales)
# Make sure we are using `DroughtData` version 2.0.0, commit 76a387a4549bb1c4e97757b37246efde33ce6d90
# install.packages("devtools")
# devtools::install_github("mountaindboz/DroughtData", ref = "76a387a4549bb1c4e97757b37246efde33ce6d90")
library(DroughtData)
library(sf)
library(here)
```

```{r check dir}
# Check if we are in the correct working directory
i_am("Sampling_Effort_DroughtData.Rmd")
```

```{r session info}
# Run session info to display package versions
devtools::session_info()
```


# Prepare Data

```{r prepare data}
# Define factor level for Region
region_lev <- c(
  "North",
  "SouthCentral",
  "Confluence",
  "Suisun Bay",
  "Suisun Marsh"
)

# Define factor level for Season
season_lev <- c(
  "Winter",
  "Spring",
  "Summer",
  "Fall"
)

# Create a nested data frame and prepare to run the summary plot function on
ndf_lt_wq <- 
  tibble(
    Parameter = c(
      "Temperature",
      "Salinity",
      "Secchi",
      "DissAmmonia",
      "DissNitrateNitrite",
      "DissOrthophos",
      "Chlorophyll"
    ),
    df_data = c(
      rep(list(raw_wq_1975_2021), 3),
      rep(list(raw_nutr_1975_2021), 3),
      list(raw_chla_1975_2021)
    )
  ) %>% 
  mutate(
    df_data_c = map2(
      df_data, 
      Parameter,
      ~ drop_na(.x, all_of(.y)) %>% 
        mutate(
          Region = factor(Region, levels = region_lev),
          Season = factor(Season, levels = season_lev)
        )
    )
  )
```


# Create Plots

```{r plot func}
# Create function for sampling effort plots
plot_samp_effort <- function(df) {
  df %>% 
    count(YearAdj, Season, Region, name = "num_samples") %>% 
    ggplot(aes(x = YearAdj, y = Region, fill = num_samples)) +
    geom_tile() +
    facet_grid(rows = vars(Season)) +
    scale_x_continuous(breaks = breaks_pretty(20), expand = expansion()) +
    scale_fill_viridis_c(name = "Number of Samples") +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
      legend.position = "top"
    )
}
```

```{r create plots}
# Create plots for each Parameter
ndf_lt_wq_plt <- ndf_lt_wq %>% mutate(plt = map(df_data_c, .f = plot_samp_effort))
```

# Sampling Effort {.tabset .tabset-pills}

```{r print plots, echo = FALSE, results = "asis", fig.width = 8, fig.height = 7}
for (i in 1:nrow(ndf_lt_wq_plt)) {
  # Create subheadings for each Parameter
  cat("## ", ndf_lt_wq_plt$Parameter[i], "\n\n")
  # Print plot
  print(ndf_lt_wq_plt$plt[[i]])
  cat("\n\n")
}
```

# Survey Summary Table

Create a table of the surveys used for each parameter in the publication and their periods of record. This table is used in the publication.

```{r summarize survey data}
df_survey_summ <- ndf_lt_wq %>%
  transmute(
    Parameter,
    df_survey = map(
      df_data_c,
      ~ group_by(.x, Source) %>%
        summarize(min_yr = min(YearAdj), max_yr = max(YearAdj))
    )
  ) %>% 
  unnest(df_survey) %>% 
  relocate(Source) %>% 
  arrange(Source, Parameter)

print(df_survey_summ, n = 40)
```

Export survey summary table to be used in the publication.

```{r export survey summary}
df_survey_summ %>% write_csv(here("outputs/survey_summary_table.csv"))
```

# Sampling Location Map

Create a two panel map of the sampling locations and regions used in the publication. One panel is for the water quality parameters and velocity stations, and the other panel is for the nutrient and chlorophyll data. This map is used in the publication.

```{r prepare station coords}
# Prepare sampling location coordinates
# Water quality measurements (water temperature, salinity, Secchi depth)
sf_wq_meas <- ndf_lt_wq %>% 
  filter(Parameter %in% c("Temperature", "Salinity", "Secchi")) %>% 
  transmute(df_coords = map(df_data_c, ~ distinct(.x, Source, Station, Latitude, Longitude))) %>% 
  unnest(df_coords) %>% 
  distinct() %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE)

# Nutrients and chlorophyll
sf_nutr_chl <- ndf_lt_wq %>% 
  filter(Parameter %in% c("DissAmmonia", "DissNitrateNitrite", "DissOrthophos", "Chlorophyll")) %>% 
  transmute(df_coords = map(df_data_c, ~ distinct(.x, Source, Station, Latitude, Longitude))) %>% 
  unnest(df_coords) %>% 
  distinct() %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE)

# Water velocity stations


# Region polygons


# Waterways basemap


# States and CA maps for inset



```

