---
title: "Temperature, Salinity, and Secchi analyses"
author: "Sam Bashevkin"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages and functions

```{r packages, message=FALSE, warning=FALSE}
require(conflicted)
require(dplyr)
# Make sure we are using `DroughtData` version 2.0.0, commit 76a387a4549bb1c4e97757b37246efde33ce6d90
# install.packages("devtools")
# devtools::install_github("mountaindboz/DroughtData", ref = "76a387a4549bb1c4e97757b37246efde33ce6d90")
require(DroughtData)
require(ggplot2)
require(patchwork)
require(knitr)
require(car)
require(multcomp)
require(emmeans)
require(stringr)
require(readr)
require(tidyr)
library(kableExtra)
library(here)

source(here("global_functions.R"))

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

Display R and package versions:

```{r}
devtools::session_info()
```

# Load data

```{r}
wq_data<-lt_avg_wq%>%
  mutate(Region=factor(Region, levels=c("Suisun Marsh", "Suisun Bay", "Confluence", "SouthCentral", "North")),
         Season=factor(Season, levels=c("Winter", "Spring", "Summer", "Fall")),
         Year_fac=factor(YearAdj),
         Drought=factor(Drought, levels=c("D", "N", "W")),
         YearAdj_s=(YearAdj-mean(YearAdj)/sd(YearAdj)),
         across(c(Secchi, Salinity), list(l=log)))
```

# Plots

Note that each facet has a different y-axis scale

## By year

### Temperature

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=YearAdj, y=Temperature, fill=Drought))+
  geom_point(shape=21, color="black")+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Temperature (째C)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

### Salinity

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=YearAdj, y=Salinity, fill=Drought))+
  geom_point(shape=21, color="black")+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Salinity")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

### Secchi

Secchi is also more regionally than seasonally variable

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=YearAdj, y=Secchi, fill=Drought))+
  geom_point(shape=21, color="black")+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Secchi depth (cm)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

# By Drought index

### Temperature

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=Drought, y=Temperature, fill=Drought))+
  geom_boxplot()+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Temperature (째C)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

### Salinity

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=Drought, y=Salinity, fill=Drought))+
  geom_boxplot()+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Salinity")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

### Secchi

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=Drought, y=Secchi, fill=Drought))+
  geom_boxplot()+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Secchi depth (cm)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

# Analyses

## Temperature

```{r}
m_t<-aov(Temperature ~ (Drought + Season + Region)^2, data=wq_data)
```

### Check assumptions:

```{r}
model_plotter(m_t, wq_data, "Temperature")
```

### Check results

```{r}
m_t_Anova<-Anova(m_t, type=3, contrasts=list(topic=contr.sum, sys=contr.sum))
m_t_Anova
```

### Partial R^2^

How much variability is explained by the Drought index?

```{r}
partial.r2(m_t_Anova)
```

### Post-hoc test

```{r}
m_t_contrasts_r <- pub_figure_plotter(df_data=wq_data, param=Temperature, y_label="Temperature (째C)", 
                                      fct_grp=Region, model=m_t, plt_title = "Water temperature")
m_t_contrasts_r$diffs

m_t_contrasts_s <- pub_figure_plotter(df_data=wq_data, param=Temperature, y_label="Temperature (째C)", 
                                      fct_grp=Season, model=m_t, plt_title = "Water temperature")
m_t_contrasts_s$diffs
```

## Salinity

Salinity data had to be log transformed to fit the assumption of normality

```{r}
m_sa<-aov(Salinity_l ~ (Drought + Season + Region)^2, data=wq_data)
```

### Check assumptions:

```{r}
model_plotter(m_sa, wq_data, "Salinity_l")
```

### Check results

```{r}
m_sa_Anova<-Anova(m_sa, type=3, contrasts=list(topic=contr.sum, sys=contr.sum))
m_sa_Anova
```

### Partial R^2^

How much variability is explained by the Drought index?

```{r}
partial.r2(m_sa_Anova)
```

### Post-hoc test

```{r}
m_sa_contrasts_r <- pub_figure_plotter(df_data=wq_data, param=Salinity, y_label="Salinity (PSU)", 
                                       fct_grp=Region, model=m_sa, log_trans = TRUE, plt_title = "Salinity")

m_sa_contrasts_r$diffs

m_sa_contrasts_s <- pub_figure_plotter(df_data=wq_data, param=Salinity, y_label="Salinity (PSU)", 
                                       fct_grp=Season, model=m_sa, log_trans = TRUE, plt_title = "Salinity")

m_sa_contrasts_s$diffs
```


## Secchi Depth

Secchi depth had to be log-transformed to comply with the assumption of normality

```{r}
m_se<-aov(Secchi_l ~ (Drought + Season + Region)^2, data=wq_data)
```

### Check assumptions:

```{r}
model_plotter(m_se, wq_data, "Secchi_l")
```

### Check results

```{r}
m_se_Anova<-Anova(m_se, type=3, contrasts=list(topic=contr.sum, sys=contr.sum))
m_se_Anova
```

### Partial R^2^

How much variability is explained by the Drought index?

```{r}
partial.r2(m_se_Anova)
```

### Post-hoc test

```{r}
m_se_contrasts_r <- pub_figure_plotter(df_data=wq_data, param=Secchi, y_label="Secchi depth (cm)", 
                                       fct_grp=Region, model=m_se, log_trans = TRUE, plt_title = "Secchi depth")

m_se_contrasts_r$diffs

m_se_contrasts_s <- pub_figure_plotter(df_data=wq_data, param=Secchi, y_label="Secchi depth (cm)", 
                                       fct_grp=Season, model=m_se, log_trans = TRUE, plt_title = "Secchi depth")

m_se_contrasts_s$diffs
```

## Secchi Depth with annual trend

Secchi depth had to be log-transformed to comply with the assumption of normality

```{r}
m_se2<-aov(Secchi_l ~ (Drought + Season + Region)^2 + YearAdj*Season*Region, data=wq_data)
```

### Check assumptions:

```{r}
model_plotter(m_se2, wq_data, "Secchi_l")
```

### Check results

```{r}
m_se2_Anova<-Anova(m_se2, type=3, contrasts=list(topic=contr.sum, sys=contr.sum))
m_se2_Anova
```

### Partial R^2^

How much variability is explained by the Drought index?

```{r}
partial.r2(m_se2_Anova)
```

### Post-hoc test

```{r}
m_se2_contrasts_r <- pub_figure_plotter(df_data=wq_data, param=Secchi, y_label="Secchi depth (cm)", 
                                        fct_grp=Region, model=m_se2, log_trans = TRUE, plt_title = "Secchi depth")

m_se2_contrasts_r$diffs

m_se2_contrasts_s <- pub_figure_plotter(df_data=wq_data, param=Secchi, y_label="Secchi depth (cm)", 
                                        fct_grp=Season, model=m_se2, log_trans = TRUE, plt_title = "Secchi depth")

m_se2_contrasts_s$diffs
```

## Save all model outputs

```{r}
# Define file paths for output and figures for the manuscript
fp_output <- here("outputs")
fp_plots <- here("plots")
```

### Anovas

```{r}
anovas <-
  bind_rows(
    mutate(as_tibble(m_t_Anova, rownames = "Parameter"), Model="Temperature"),
    mutate(as_tibble(m_sa_Anova, rownames = "Parameter"), Model="Salinity"),
    mutate(as_tibble(m_se2_Anova, rownames = "Parameter"), Model="Secchi depth")
  ) %>%
  mutate(
    `Pr(>F)`=if_else(
      `Pr(>F)`<0.001, 
      "< 0.001", 
      as.character(formatC(signif(`Pr(>F)`, 4), digits = 4, format = "fg", flag = "#"))
    ),
    across(c(`Sum Sq`, `F value`), ~formatC(signif(.x, 4), digits = 4, format = "fg", flag = "#"))
  ) %>%
  relocate(Model)
  
kbl(anovas) %>% kable_styling(fixed_thead = TRUE)

# Excel is the worst, I had to add letters to these numbers so it wouldn't re-round them
anovas %>% 
  mutate(
    across(
      c(`Sum Sq`, `F value`, `Pr(>F)`), 
      ~paste0(ifelse(str_detect(.x, fixed("NA")) | is.na(.x), "", .x), "ZZZ")
    )
  ) %>% 
  write_csv(file.path(fp_output, "TempSalSecchi_anovas.csv"))
```

### Post-hoc tests

```{r}
tukeys <- 
  bind_rows(
    mutate(as_tibble(m_t_contrasts_r$contrasts), Model="Temperature", `Contrast type`="region"),
    mutate(as_tibble(m_t_contrasts_s$contrasts), Model="Temperature", `Contrast type`="season"),
    mutate(as_tibble(m_sa_contrasts_r$contrasts), Model="Salinity", `Contrast type`="region"),
    mutate(as_tibble(m_sa_contrasts_s$contrasts), Model="Salinity", `Contrast type`="season"),
    mutate(as_tibble(m_se2_contrasts_r$contrasts), Model="Secchi depth", `Contrast type`="region"),
    mutate(as_tibble(m_se2_contrasts_s$contrasts), Model="Secchi depth", `Contrast type`="season")
  ) %>%
  rename(
    `p-value`=p.value, 
    `t-ratio`=t.ratio
  ) %>%
  mutate(
    `p-value`=if_else(
      `p-value`<0.001, 
      "< 0.001", 
      as.character(formatC(signif(`p-value`, 4), digits=4, format="fg", flag="#"))
    ),
    Group=if_else(is.na(Region), as.character(Season), as.character(Region)),
    across(c(estimate, SE, `t-ratio`), ~formatC(signif(.x, 4), digits=4, format="fg", flag="#"))
  ) %>%
  select(
    Model,
    `Contrast type`,
    Group, 
    Contrast=contrast,
    Estimate=estimate, 
    SE,
    DF=df, 
    `t-ratio`, 
    `p-value`
  )
  
kbl(tukeys) %>% kable_styling(fixed_thead = TRUE)

# Excel is the worst, I had to add letters to these numbers so it wouldn't re-round them
tukeys %>% 
  mutate(
    across(
      c(Estimate, SE, `t-ratio`, `p-value`), 
      ~paste0(ifelse(str_detect(.x, fixed("NA")) | is.na(.x), "", .x), "ZZZ")
    )
  ) %>% 
  write_csv(file.path(fp_output, "TempSalSecchi_tukeys.csv"))
```

### Manuscript Figures

Combine all figures of model results for the water quality parameters into larger figures for the manuscript, one for the regional analyses and one for the seasonal analyses.

```{r}
# Region plots
plt_wq_region <- m_t_contrasts_r$plt + m_sa_contrasts_r$plt + m_se2_contrasts_r$plt +
  plot_layout(ncol = 1)

ggsave(
  file.path(fp_plots, "water_quality_region.jpg"),
  plot = plt_wq_region,
  width = 8,
  height = 9,
  units = "in",
  dpi = 300
)

# Season plots
plt_wq_season <- m_t_contrasts_s$plt + m_sa_contrasts_s$plt + m_se2_contrasts_s$plt +
  plot_layout(ncol = 1)

ggsave(
  file.path(fp_plots, "water_quality_season.jpg"),
  plot = plt_wq_season,
  width = 8,
  height = 9,
  units = "in",
  dpi = 300
)
```

