---
title: "Temperature, Salinity, and Secchi analyses"
author: "Sam Bashevkin"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    )
    })
    rmarkdown::render(
      input,
      output_dir = 'docs',
      knit_root_dir = "../",
      envir = globalenv()
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE, warning=FALSE}
require(dplyr)
require(DroughtData)
require(ggplot2)
require(patchwork)
require(knitr)
require(car)
require(multcomp)
require(emmeans)
require(stringr)
require(readr)
```

# Functions

```{r}
model_plotter<-function(model, data, parameter){
  data<-data%>%
    filter(!is.na(.data[[parameter]]))%>%
    mutate(Residuals=resid(model),
           Fitted=predict(model))
  
  units<-case_when(
    parameter=="Temperature" ~ " (°C)",
    parameter=="Salinity_l" ~ " (log)",
    parameter=="Secchi_l" ~ " (log)" 
  )
  parameter_label<-case_when(parameter=="Secchi_l" ~ "secchi depth", 
                             parameter=="Salinity_l" ~ "salinity",
                             parameter=="Temperature" ~ "temperature")
  
  p_hist<-ggplot(data, aes(x=Residuals))+
    geom_histogram()+
    xlab(paste0("Residuals", units))+
    theme_bw()
  
  p_res_fit<-ggplot(data, aes(x=Residuals, y=Fitted))+
    geom_point()+
    ylab(paste0("Predicted ", parameter_label, units))+
    xlab(paste0("Residuals", units))+
    theme_bw()
  
  p_obs_fit<-ggplot(data, aes(x=.data[[parameter]], y=Fitted))+
    geom_point()+
    geom_abline(slope=1, intercept=0, color="red")+
    ylab(paste0("Predicted ", parameter_label, units))+
    xlab(paste0("Observed ", parameter_label, units))+
    theme_bw()
  
  out<-(p_hist+plot_layout(ncol=1))+(p_res_fit+p_obs_fit+plot_layout(ncol=2))+plot_layout(nrow=2, widths=c(1, 0.5, 0.5))
  
  return(out)
}

result_plotter<-function(contrast, variable, xlabel){
  contrasts<-plot(contrast$emmeans, comparisons = T, by=variable, plotit=F)%>%
    mutate(across(c(lcmpl, rcmpl), ~ if_else(is.na(.x), the.emmean, .x)))
  
  ggplot(contrasts, aes(y=Drought, x=the.emmean))+
    geom_pointrange(aes(xmin=lcmpl, xmax=rcmpl))+
    facet_wrap(~.data[[variable]], scales="free_x", ncol=1)+
    xlab(xlabel)+
    theme_bw()
}

partial.r2<-function(ANOVA, factor){
  r2<-ANOVA[factor, "Sum Sq"]/(ANOVA[factor, "Sum Sq"]+ ANOVA["Residuals", "Sum Sq"])
  return(r2)
}
```

# Load data

```{r}
wq_data<-lt_avg_wq%>%
  mutate(Region=factor(Region, levels=c("Suisun Marsh", "Suisun Bay", "Confluence", "SouthCentral", "North")),
         Season=factor(Season, levels=c("Winter", "Spring", "Summer", "Fall")),
         Year_fac=factor(YearAdj),
         Drought=factor(Drought, levels=c("D", "N", "W")),
         YearAdj_s=(YearAdj-mean(YearAdj)/sd(YearAdj)),
         across(c(Secchi, Salinity), list(l=log)))
```

# Plots

Note that each facet has a different y-axis scale

## By year

### Temperature

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=YearAdj, y=Temperature, fill=Drought))+
  geom_point(shape=21, color="black")+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Temperature (°C)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

### Salinity

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=YearAdj, y=Salinity, fill=Drought))+
  geom_point(shape=21, color="black")+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Salinity")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

### Secchi

Secchi is also more regionally than seasonally variable

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=YearAdj, y=Secchi, fill=Drought))+
  geom_point(shape=21, color="black")+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Secchi depth (cm)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

# By Drought index

### Temperature

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=Drought, y=Temperature, fill=Drought))+
  geom_boxplot()+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Temperature (°C)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

### Salinity

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=Drought, y=Salinity, fill=Drought))+
  geom_boxplot()+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Salinity")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

### Secchi

```{r, fig.width=12, fig.height=12}
ggplot(wq_data, aes(x=Drought, y=Secchi, fill=Drought))+
  geom_boxplot()+
  facet_wrap(Season~Region, scales="free_y")+
  drt_color_pal_drought()+
  ylab("Secchi depth (cm)")+
  theme_bw()+
  theme(axis.text.x=element_text(angle=45, hjust=1))
```

# Analyses

## Temperature

```{r}
m_t<-aov(Temperature ~ (Drought + Season + Region)^2, data=wq_data)
```

### Check assumptions:

```{r}
model_plotter(m_t, wq_data, "Temperature")
```

### Check results

```{r}
m_t_Anova<-Anova(m_t, type=2)
m_t_Anova
```

### Partial R^2^

How much variability is explained by the Drought index?

```{r}
partial.r2(m_t_Anova, "Drought")
```

### Post-hoc test

```{r}
m_t_contrasts_r <- emmeans(m_t, pairwise ~ Drought*Region)

m_t_contrasts_s <- emmeans(m_t, pairwise ~ Drought*Season)
```

Option 1: non-overlapping error bars = significant difference
```{r, fig.width=4, fig.height=6}
# Region
result_plotter(m_t_contrasts_r, "Region", "Temperature (°C)")

# Season
result_plotter(m_t_contrasts_s, "Season", "Temperature (°C)")
```

## Salinity

Salinity data had to be log transformed to fit the assumption of normality

```{r}
m_sa<-aov(Salinity_l ~ (Drought + Season + Region)^2, data=wq_data)
```

### Check assumptions:

```{r}
model_plotter(m_sa, wq_data, "Salinity_l")
```

### Check results

```{r}
m_sa_Anova<-Anova(m_sa, type=2)
m_sa_Anova
```

### Partial R^2^

How much variability is explained by the Drought index?

```{r}
partial.r2(m_sa_Anova, "Drought")
```

### Post-hoc test

```{r}
m_sa_contrasts_r <- emmeans(m_sa, pairwise ~ Drought*Region)

m_sa_contrasts_s <- emmeans(m_sa, pairwise ~ Drought*Season)
```

Option 1: non-overlapping error bars = significant difference
```{r, fig.width=4, fig.height=6}
# Region
result_plotter(m_sa_contrasts_r, "Region", "Salinity")

# Season
result_plotter(m_sa_contrasts_s, "Season", "Salinity")
```

## Secchi Depth

Secchi depth had to be log-transformed to comply with the assumption of normality

```{r}
m_se<-aov(Secchi_l ~ (Drought + Season + Region)^2, data=wq_data)
```

### Check assumptions:

```{r}
model_plotter(m_se, wq_data, "Secchi_l")
```

### Check results

```{r}
m_se_Anova<-Anova(m_se, type=2)
m_se_Anova
```

### Partial R^2^

How much variability is explained by the Drought index?

```{r}
partial.r2(m_se_Anova, "Drought")
```

### Post-hoc test

```{r}
m_se_contrasts_r <- emmeans(m_se, pairwise ~ Drought*Region)

m_se_contrasts_s <- emmeans(m_se, pairwise ~ Drought*Season)
```

Option 1: non-overlapping error bars = significant difference
```{r, fig.width=4, fig.height=6}
# Region
result_plotter(m_se_contrasts_r, "Region", "Secchi depth (log)")

# Season
result_plotter(m_se_contrasts_s, "Season", "Secchi depth (log)")
```

## Secchi Depth with annual trend

Secchi depth had to be log-transformed to comply with the assumption of normality

```{r}
m_se2<-aov(Secchi_l ~ (Drought + Season + Region)^2 + YearAdj*Season*Region, data=wq_data)
```

### Check assumptions:

```{r}
model_plotter(m_se2, wq_data, "Secchi_l")
```

### Check results

```{r}
m_se2_Anova<-Anova(m_se2, type=2)
m_se2_Anova
```

### Partial R^2^

How much variability is explained by the Drought index?

```{r}
partial.r2(m_se2_Anova, "Drought")
```

### Post-hoc test

```{r}
m_se2_contrasts_r <- emmeans(m_se2, pairwise ~ Drought*Region)

m_se2_contrasts_s <- emmeans(m_se2, pairwise ~ Drought*Season)
```

Option 1: non-overlapping error bars = significant difference
```{r, fig.width=4, fig.height=6}
# Region
result_plotter(m_se2_contrasts_r, "Region", "Secchi depth (log)")

# Season
result_plotter(m_se2_contrasts_s, "Season", "Secchi depth (log)")
```

## Save all model outputs

Anovas
```{r}
anovas<-bind_rows(
  mutate(as_tibble(m_t_Anova, rownames = "Parameter"), model="Temperature"),
  mutate(as_tibble(m_sa_Anova, rownames = "Parameter"), model="Salinity"),
  mutate(as_tibble(m_se_Anova, rownames = "Parameter"), model="Secchi depth")
)%>%
  mutate(`Pr(>F)`=if_else(`Pr(>F)`<0.001, "< 0.001", as.character(round(`Pr(>F)`, 4))))%>%
  relocate(model)%>%
  write_csv(file.path("outputs", "TempSalSecchi_anovas.csv"))

print(anovas, n=1000)
```

Post-hoc tests
```{r}
tukeys<-bind_rows(
  mutate(as_tibble(m_t_contrasts_r$contrasts), model="Temperature", `contrast type`="region"),
  mutate(as_tibble(m_t_contrasts_s$contrasts), model="Temperature", `contrast type`="season"),
  mutate(as_tibble(m_sa_contrasts_r$contrasts), model="Salinity", `contrast type`="region"),
  mutate(as_tibble(m_sa_contrasts_s$contrasts), model="Salinity", `contrast type`="season"),
  mutate(as_tibble(m_se_contrasts_r$contrasts), model="Secchi depth", `contrast type`="region"),
  mutate(as_tibble(m_se_contrasts_s$contrasts), model="Secchi depth", `contrast type`="season"))%>%
  rename(`p-value`=p.value, `t-ratio`=t.ratio)%>%
  mutate(`p-value`=if_else(`p-value`<0.001, "< 0.001", as.character(round(`p-value`, 4))))%>%
  relocate(model, `contrast type`)%>%
  write_csv(file.path("outputs", "TempSalSecchi_tukeys.csv"))

print(anovas, n=1000)
```
