---
title: "Fine scale velocity exploration for water quality drought synthesis"
author: "Elizabeth Stumpner"
date: "2022-12-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```


```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(ggbeeswarm)
library(DroughtData)
library(sf)
library(mapview)
library(stats)
```
# Overview

This is the velocity analysis for the Drought paper. The weekly  velocity records are calculated from instantanous (15-min) records at four USGS flow stations. Stations include Cache at Ryer Island Ferry, San Joaquin River at Jersey Point, Middle River, and Old River at Bacon Island (see map below). The weekly velocity data is combined with dayflow parameters and drought classification in the DroughtData R package.

```{r}
vel_weekly <- read_rds("velocity.rds")

lat_long <- read_csv('vel_coord.csv')
```

## Station map 

Here is a map of the selected USGS stations flow stations in the Delta - from north to south the stations include Cache Slough (USGS 11455350 - inactive RYI station, 11455385 - active RYF station), Jersey Point (11337190), Old River at Bacon Island (11313405), and Middle River (11312676)

```{r}
lat_long_sf <- st_as_sf(lat_long, coords = c('LongitudeMeasure', 'LatitudeMeasure'), crs = 4326)

mapview(lat_long_sf, layer.name = "Velocity", cex=10, label = 'site')

```

## Figure 3 - velocity & delta outflow by station 

Next up is exploring the relationship with delta outflow, and mean net velocity and maximum absolute tidal velocity at four USGS flow stations. Net velocities only increase considerably during periods of high outflow in Cache Slough during wet years. Higher net velocities in Cache Slough reflect periods of Yolo Bypass inundation - when water management activities divert floodwaters from the mainstem Sacramento River. Tidal velocities generally fall within a similar range across all water year types but exhibit a downward trend during periods of high outflow. The negative velocity direction is consistent with periods of higher freshwater flow that mute tidal forcing.

```{r}
fig_3a <- vel_weekly %>%
  ggplot(aes(x=Log_Outflow_m, y=mean_net_vel_m, color = YearType, shape = net_sign))+
  drt_color_pal_yrtype(aes_type = 'color') +
  geom_point(alpha=0.4, cex = 2.5)+
  facet_wrap(~station, scales = "free") +
  theme (text = element_text (size = 15)) +
  geom_hline(yintercept=0, linetype='solid', col = 'black') +
  labs(y = expression(paste("mean net vel. (", m," ",s^-1, ")", sep="")), x = expression(paste("log(outflow) (", m^3, " ",s^-1, ")", sep="")), title = "A", color = "Year Type", shape = "Sign") +
  scale_shape(labels = c("negative", "positive")) +
  guides(color = guide_legend('Water year index'))

fig_3a

fig_3b <- vel_weekly %>%
  ggplot(aes(x=Log_Outflow_m, y=max_abs_tidal_vel_m, color=YearType, shape = tide_sign))+
  geom_point(alpha=0.4, cex = 2.5)+
  drt_color_pal_yrtype(aes_type = 'color') +
  facet_wrap(~station, scales = "free") +
  theme (text = element_text (size = 15)) +
  #geom_hline(yintercept=0, linetype='solid', col = 'black') +
  labs(y = expression(paste("max. abs. tidal vel. (", m," ",s^-1, ")", sep="")), x = expression(paste("log(outflow) (", m^3, " ",s^-1, ")", sep="")), title = "B", color = "Year Type", shape = "Sign") +
  scale_shape(labels = c("negative", "positive"))+
  guides(color = guide_legend('Water year index'))

fig_3b
```


## Figure S3 - Delta outflow 
To start, delta outflow is similar at the start of most water years except for wet years when outflow is greater at the beginning. Smoothing the data of each water year type shows outflow 'peaks' earlier in the water year during critical and drier years relative to wetter water years. This pattern likely reflects water management activities to increase reservoir storage.

```{r}
vel_weekly %>%
  ggplot(aes(x=wtr_day, y=Log_Outflow_m, color = YearType))+
  drt_color_pal_yrtype(aes_type = 'color') +
  geom_smooth(alpha=1)+
  theme(text = element_text(size = 22)) +
  theme_bw()+
  guides(color = guide_legend('Water year index')) +
  labs(x = "Water day of year", y = expression(paste("log(outflow) (", m^3," ", s^-1, ")", sep = "")))
```

## Figure S4 - mean net velocity and maximum absolute tidal velocity with WY index across stations

Across all water years, mean net velocity is most always positive at Cache and Jersey but only positive in the wet seasons of wet years at the Middle and Old River stations. For much of the year, the export pumps create net landward velocities in Middle and Old Rivers that is demonstrated by negative velocities. Only during the wet season of wet years does increased freshwater flow overcome the effects of the pumps and create net positive velocities. 
```{r}
# Fig. S4A
vel_weekly %>%
  ggplot(aes(x=wtr_day, y=mean_net_vel_m, color = YearType)) +
  geom_smooth() +
  drt_color_pal_yrtype(aes = 'color') +
  facet_wrap(~station, scales = "fixed")+
  geom_hline(yintercept=0, linetype='solid', col = 'black') +
  labs(x = "Water day of year", y = expression(paste("mean net vel. (", m," ",s^-1, ")", sep="")), title = "A") +
  theme (text = element_text (size = 12)) +
  guides(color = guide_legend('Water year index'))

# Fig. S4B

vel_weekly %>%
  ggplot(aes(x=wtr_day, y=max_abs_tidal_vel_m, color = YearType)) +
  geom_smooth() +
  drt_color_pal_yrtype(aes = 'color') +
  facet_wrap(~station, scales = "free")+
  labs(x = "Water day of year", y = expression(paste("max. abs. tidal vel. (", m," ",s^-1, ")", sep="")), title = "B") +
  scale_y_continuous(n.breaks = 3) +
  theme (text = element_text (size = 12)) +
  guides(color = guide_legend('Water year index'))
```

## Compare 2020 + 2021 to other years - WQ White Paper

### Mean net velocity and maximum absolute velocity across stations
The 2020 and 2021 net velocities compare well to previous drought years. Breaking the data down by station show the Cache and Jersey net velocities are most always seaward and Middle and Old Rivers most always landward, except during wet periods of wet years.

The max. abs. tidal velocities in 2020 and 2021 are nearly identical to dry and critical years years except at Cache where they're slightly lower. With data faceted by station, a slight reduction in max. abs. tidal vel. range at Cache is observed in 2020 relative to other years.
```{r}

#zoom in
ggplot(vel_weekly, aes(x=YearType_20_21, y=mean_net_vel_m, fill=YearType))+
  geom_boxplot()+
  facet_wrap(~station) +
  drt_color_pal_yrtype()+
  labs(x = "Water year index", y = expression(paste("mean net vel. (", m," ",s^-1, ")", sep="")), title = "A. mean net vel. across stations (2007 - 2021) - restricted y axis") +
  guides(fill = guide_legend('Water year index')) +
  ylim(c(-0.25, 0.25))+
  theme_bw()+
  geom_hline(yintercept=0, linetype='solid', col = 'black') +
  theme(axis.text.x = element_text(angle = 90)) 
  
ggplot(vel_weekly, aes(x=YearType_20_21, y=max_abs_tidal_vel_m, fill=YearType))+
  geom_boxplot()+
  facet_wrap(~station)+
  drt_color_pal_yrtype()+
  labs(x = "Water year index", y = expression(paste("max. abs. tidal vel. (", m," ",s^-1, ")", sep="")), title = "B. max. abs. tidal vel. across stations (2007 - 2021)") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90)) +
  guides(fill = guide_legend('Water year index'))
```



