---
title: "Long-term WQ: Sampling Effort"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Look at the sampling effort for the various parameters to be included in the long-term WQ publication. This will help us decide how to structure our ANOVA models. Sampling effort is visualized with tile plots of samples collected by region and season and maps displaying spatial coverage. This document also contains a summary table of the surveys, parameters, and periods of record for the publication.


# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(sf)
library(qs)
library(here)
library(conflicted)
```

```{r declare conflict pref}
# Declare package conflict preferences
conflicts_prefer(dplyr::filter())
```

```{r check dir}
# Check if we are in the correct working directory
i_am("data_processing/Sampling_Effort_DroughtData.Rmd")
```

```{r session info}
# Run session info to display package versions
devtools::session_info()
```

# Import and Prepare Data

```{r import data}
# Define file paths for raw data for WQ measurements, nutrients, and chlorophyll
fp_raw_data <- dir(path = here("data"), pattern = "^raw.+qs$", full.names = TRUE)

# Import and combine raw data
df_raw_data <- map(fp_raw_data, qread) %>% list_rbind()
```

```{r prepare data}
# Define factor level for Region
region_lev <- c(
  "North",
  "SouthCentral",
  "Confluence",
  "Suisun Bay",
  "Suisun Marsh"
)

# Define factor level for Season
season_lev <- c(
  "Winter",
  "Spring",
  "Summer",
  "Fall"
)

# Define factor level for Parameter
param_lev <- c(
  "Temperature",
  "Salinity",
  "Secchi",
  "DissAmmonia",
  "DissNitrateNitrite",
  "DissOrthophos",
  "Chlorophyll"
)

# Prepare raw data to run the summary plot function on
df_raw_data_c <- df_raw_data %>% 
  mutate(
    Region = factor(Region, levels = region_lev),
    Season = factor(Season, levels = season_lev),
    Parameter = factor(Parameter, levels = param_lev)
  )
```

# Create Plots

```{r plot func}
# Create function for sampling effort plots
plot_samp_effort <- function(df) {
  df %>% 
    count(YearAdj, Season, Region, name = "num_samples") %>% 
    ggplot(aes(x = YearAdj, y = Region, fill = num_samples)) +
    geom_tile() +
    facet_grid(rows = vars(Season)) +
    scale_x_continuous(breaks = breaks_pretty(20), expand = expansion()) +
    scale_fill_viridis_c(name = "Number of Samples") +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
      legend.position = "top"
    )
}
```

```{r create plots}
# Create plots for each Parameter
ndf_se_plt <- df_raw_data_c %>% 
  nest(.by = Parameter, .key = "df_data") %>% 
  mutate(plt = map(df_data, .f = plot_samp_effort)) %>% 
  arrange(Parameter)
```

# Sampling Effort {.tabset .tabset-pills}

```{r print plots, echo = FALSE, results = "asis", fig.width = 8, fig.height = 7}
for (i in 1:nrow(ndf_se_plt)) {
  # Create subheadings for each Parameter
  cat("## ", ndf_se_plt$Parameter[i], "\n\n")
  # Print plot
  print(ndf_se_plt$plt[[i]])
  cat("\n\n")
}
```

# Regional Coverage

Placeholder

# Survey Summary Table

Create a table of the surveys used for each parameter in the publication and their periods of record. This table is used in the publication.

```{r summarize survey data}
df_survey_summ <- df_raw_data_c %>%
  summarize(
    min_yr = min(YearAdj), 
    max_yr = max(YearAdj),
    .by = c(Source, Parameter)
  ) %>% 
  arrange(Source, Parameter)

print(df_survey_summ, n = 50)
```

Export survey summary table to be used in the publication.

```{r export survey summary}
df_survey_summ %>% write_csv(here("outputs/survey_summary_table.csv"))
```

