---
title: "Inflow and outflow"
author: "Sam Bashevkin"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = here::here("docs"),
      envir = globalenv()
    )
    })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
```

Display R and package versions:

```{r}
devtools::session_info()
```

# Data prep

```{r}
# Import Delta Inflow and Outflow data from Dayflow
dayflow <- readRDS(here("data/processed/hydrology/dayflow.rds"))

# Prepare Dayflow data for plots
dayflow_c <- dayflow %>%
  mutate(
    Period_year_cat = factor(Period_year_cat, levels = c("1", "2", "3+")),
    Drought = factor(
      Drought, 
      levels = c("D", "N", "W"), 
      labels = c("Drought", "Neutral", "Wet")
    )
  )
```

# Inflow plots

Total inflow seasonal hydrograph by year

```{r, warning = FALSE}
dayflow_c %>% 
  ggplot(aes(x = WYday, y = InflowTotal, color = Period_year_cat, group = WaterYear)) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 25)) +
  facet_grid(rows = vars(Drought), scales = "free_y") +
  scale_color_viridis_d(name = "Period-year") +
  coord_cartesian(ylim = c(0, NA), expand = FALSE) +
  theme_bw()
```

Total inflow seasonal hydrograph by Period_year

```{r, warning = FALSE}
dayflow_c %>%
  ggplot(aes(x = WYday, y = InflowTotal, color = Period_year_cat, group = Period_year_cat)) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 25)) +
  facet_grid(rows = vars(Drought), scales = "free_y") +
  scale_color_viridis_d(name = "Period-year") +
  scale_x_continuous(
    name = "Day of Water Year",
    breaks = seq.int(0, 360, by = 60),
    expand = expansion(mult = 0.005)
  ) +
  scale_y_continuous(limits = c(0, NA), expand = expansion()) +
  theme_bw()
```

Total inflow seasonal hydrograph by Period_year, scaled by summed inflow over the year

```{r, warning = FALSE}
dayflow_c %>% 
  ggplot(aes(x = WYday, y = InflowTotal / InflowTotal_sum, color = Period_year_cat)) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 25)) +
  facet_grid(rows = vars(Drought)) +
  scale_color_viridis_d(name = "Period-year") +
  scale_x_continuous(
    name = "Day of Water Year",
    breaks = seq.int(0, 360, by = 60),
    expand = expansion(mult = 0.005)
  ) +
  scale_y_continuous(limits = c(0, NA), expand = expansion()) +
  theme_bw()
```

Total cumulative inflow seasonal hydrograph by Period_year

```{r}
plt_inflow_wyday <- dayflow_c %>% 
  ggplot(
    aes(
      x = WYday, 
      y = InflowTotal_cum / InflowTotal_sum, 
      color = Period_year_cat, 
      group = WaterYear, 
      linetype = Period_year_cat
    )
  ) +
  geom_line() +
  facet_grid(rows = vars(Drought), scales = "free_y") +
  scale_color_viridis_d(name = "Period-year") +
  scale_linetype(name = "Period-year") +
  scale_x_continuous(
    name = "Day of Water Year",
    breaks = seq.int(0, 360, by = 60),
    expand = expansion(mult = 0.005)
  ) +
  scale_y_continuous(
    name = "Cumulative proportional inflow",
    limits = c(0, NA), 
    expand = expansion()
  ) +
  theme_bw() +
  theme(panel.spacing = unit(1, "lines"))

plt_inflow_wyday
```

This shows most clearly how in dry years, after 2+ years of a drought, water managers are more conservative early in the water year, saving up whatever rainfall they can. But in wet years after 2+ years of a wet period, the reservoirs are full and much more water is released earlier in the water year. Neutral years show no pattern, as expected since they represent flip-flopping among year types.

Same plot as above, but smoothed for each Period_year and drought category combo

```{r, warning = FALSE}
dayflow_c %>% 
  ggplot(
    aes(
      x = WYday, 
      y = InflowTotal_cum / InflowTotal_sum, 
      color = Period_year_cat, 
      fill = Period_year_cat, 
      linetype = Period_year_cat
    )
  ) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 25), alpha = 0.5) +
  facet_grid(rows = vars(Drought), scales = "free_y") +
  scale_color_viridis_d(aesthetics = c("color", "fill"), name = "Period-year") +
  scale_linetype(name = "Period-year") +
   scale_x_continuous(
    name = "Day of Water Year",
    breaks = seq.int(0, 360, by = 60),
    expand = expansion(mult = 0.005)
  ) +
  scale_y_continuous(
    name = "Cumulative proportional inflow",
    limits = c(0, NA), 
    expand = expansion()
  ) +
  theme_bw()
```

# Outflow plots

Total cumulative outflow seasonal hydrograph by Period_year

```{r}
plt_outflow_wyday <- dayflow_c %>% 
  ggplot(
    aes(
      x = WYday, 
      y = Outflow_cum / Outflow_sum, 
      color = Period_year_cat, 
      group = WaterYear, 
      linetype = Period_year_cat
    )
  ) +
  geom_line() +
  facet_grid(rows = vars(Drought), scales = "free_y") +
  ylab("Cumulative proportional outflow") +
  scale_color_viridis_d(name = "Period-year") +
  scale_linetype(name = "Period-year") +
   scale_x_continuous(
    name = "Day of Water Year",
    breaks = seq.int(0, 360, by = 60),
    expand = expansion(mult = 0.005)
  ) +
  scale_y_continuous(
    name = "Cumulative proportional outflow",
    limits = c(0, NA), 
    expand = expansion()
  ) +
  theme_bw() +
  theme(panel.spacing = unit(1, "lines"))

plt_outflow_wyday
```

# Export Figures

```{r}
# Define file path for manuscript figures
fp_plots <- here("results/figures")

# Cumulative inflow figure
ggsave(
  file.path(fp_plots, "inflow_wyday.jpg"), 
  plot = plt_inflow_wyday, 
  width = 6, 
  height = 4, 
  units = "in",
  dpi = 300
)

# Cumulative outflow figure
ggsave(
  file.path(fp_plots, "outflow_wyday.jpg"), 
  plot = plt_outflow_wyday, 
  width = 6, 
  height = 4, 
  units = "in",
  dpi = 300
)
```

